---
title: 'SQL Practice: Advanced problems'
output: html_document
---

```{r setup, include=FALSE}
library(RSQLite)
library(tidyverse)
library(lubridate)


#con <- dbConnect(RSQLite::SQLite(), "db/Northwind_large.sqlite")
#min_db <- dbConnect(RSQLite::SQLite(), "db/Northwind.sqlite")
con <- dbConnect(RSQLite::SQLite(), "db/edit_Northwind.sqlite")

knitr::opts_chunk$set(echo = TRUE, connection = "con")
```

## Connect to DB and overview tables
```{r}
dbListTables(con)
```

### Q32: High-value customers  

We want to send all of our high-value customers a special VIP gift. We're defining high-value customers as those who've made at least 1 order with a total value (not including the discount) equal to $10,000 or more. We only want to consider orders made in the year 2016.

Look into **OrderDetail** table
```{sql}
SELECT * FROM 'OrderDetail'
```
What is the length of **OrderDetails**?
```{sql}
SELECT COUNT(*) AS 'Number Of Orders'
FROM 'OrderDetail'
```

Look into **Orders** table
```{sql}
SELECT * FROM 'Orders'
```

Look into **Customer** table
```{sql}
SELECT * FROM 'Customer'
```

Step 1. Show order details for order No 10248
```{sql}
SELECT *
FROM 'OrderDetail'
WHERE OrderId = 10248
```

Step 2. Count total value (UnitPrice*Quantity) for order No 10248
```{sql}
SELECT OrderId, SUM(UnitPrice*Quantity) AS 'TotalValue'
FROM 'OrderDetail'
WHERE OrderId = 10248
```
Test results:  14\*12+9.8\*10+34.8*5 = 440  
  
Step 3. Count total value for orders:
```{sql}
SELECT OrderId, SUM(UnitPrice*Quantity) AS 'TotalValue'
FROM 'OrderDetail'
--WHERE OrderId IN (10248, 10250)
GROUP BY OrderId
```

### Now let's solve our problem in tidyverse manere:)  
  
Let's loads R object with sql-tables:  
```{r}
load("db/small_customer_table.Rdata")
load("db/small_order_details_table.Rdata")
load("db/small_orders_table.Rdata")
```

See orders table:
```{r}
small_orders
```

Next, let's see number of orders by years
```{r}
small_orders %>% 
  count(Year = year(OrderDate))
```
See order_details table:
```{r}
small_order_details
```
See customers table:
```{r}
small_customer
```

Repeat our task: We want to send all of our high-value customers a special VIP gift. We're defining high-value customers as those who've made at least 1 order with a total value (not including the discount) equal to $10,000 or more. We only want to consider orders made in the year 2016.  
  
Next, view total orders for order No 10248:
```{r}
inner_join(select(small_order_details, 1:5),
           select(small_orders, Id, CustomerId, OrderDate) %>% filter(., Id == 10248),
           by = c("OrderId" = "Id"))
```
Join customer's name:
```{r}
inner_join(select(small_order_details, 1:5),
           select(small_orders, Id, CustomerId, OrderDate) %>% filter(Id %in% c(10248, 10252)),
           by = c("OrderId" = "Id")) %>% 
  inner_join(select(small_customer, Id, CompanyName), by = c("CustomerId" = "Id"))
```
Count Total Value:
```{r}
inner_join(select(small_order_details, 1:5),
           select(small_orders, Id, CustomerId, OrderDate) %>% filter(Id %in% c(10248, 10252)),
           by = c("OrderId" = "Id")) %>% 
  inner_join(select(small_customer, Id, CompanyName), by = c("CustomerId" = "Id")) %>% 
  group_by(CustomerId) %>% 
  summarise(TotalValue = sum(UnitPrice * Quantity))
```
Final desision:
```{r}
inner_join(select(small_order_details, 1:5),
           select(small_orders, Id, CustomerId, OrderDate) %>% filter(year(OrderDate) == 2016),
           by = c("OrderId" = "Id")) %>% 
  inner_join(select(small_customer, Id, CompanyName), by = c("CustomerId" = "Id")) %>% 
  group_by(CustomerId, OrderId) %>% #CompanyName
  summarise(TotalValue = sum(UnitPrice * Quantity)) %>% 
  filter(TotalValue > 10000) %>% 
  arrange(desc(TotalValue))
```
My decision for SQL:
```{sql}
SELECT Orders.CustomerId, Customer.CompanyName, OrderDetail.OrderId, SUM(OrderDetail.UnitPrice*OrderDetail.Quantity) AS 'TotalValue'
FROM 'OrderDetail'
JOIN 'Orders' ON OrderDetail.OrderId = Orders.Id 
JOIN 'Customer' ON Customer.Id = Orders.CustomerId
WHERE strftime('%Y', Orders.OrderDate) = '2016'
GROUP BY CustomerId, OrderId
HAVING SUM(OrderDetail.UnitPrice*OrderDetail.Quantity) > 10000
ORDER BY TotalValue DESC
```

### Q33: High-value customers - total orders  
  
The manager has changed his mind. Instead of requiring that customers have at least one individual orders totaling $10,000 or more, he wants to
define high-value customers as those who have orders totaling $15,000 or more in 2016. How would you change the answer to the problem above?
```{sql}
SELECT Orders.CustomerId, Customer.CompanyName, OrderDetail.OrderId, SUM(OrderDetail.UnitPrice*OrderDetail.Quantity) AS 'TotalValue'
FROM 'OrderDetail'
JOIN 'Orders' ON OrderDetail.OrderId = Orders.Id 
JOIN 'Customer' ON Customer.Id = Orders.CustomerId
WHERE strftime('%Y', Orders.OrderDate) = '2016'
GROUP BY CustomerId
HAVING SUM(OrderDetail.UnitPrice*OrderDetail.Quantity) > 15000
ORDER BY TotalValue DESC
```

### Q34: High-value customers - with discount  
  
Change the above query to use the discount when calculating high-value customers. Order by the total amount which includes the discount. 
```{sql}
SELECT Orders.CustomerId, Customer.CompanyName, OrderDetail.OrderId,
       SUM(OrderDetail.UnitPrice*OrderDetail.Quantity*(1 - OrderDetail.Discount)) AS 'TotalValueWithDiscount'
FROM 'OrderDetail'
JOIN 'Orders' ON OrderDetail.OrderId = Orders.Id 
JOIN 'Customer' ON Customer.Id = Orders.CustomerId
WHERE strftime('%Y', Orders.OrderDate) = '2016'
GROUP BY CustomerId
HAVING SUM(OrderDetail.UnitPrice*OrderDetail.Quantity*(1 - OrderDetail.Discount)) > 10000
ORDER BY TotalValueWithDiscount DESC
```

### Q35: Month-end orders

At the end of the month, salespeople are likely to try much harder to get orders, to meet their month-end quotas. Show all orders made on the last day of the month. Order by EmployeeID and OrderID.  
  
First, show Orders table:
```{sql}
SELECT * FROM 'Orders'
```
Final answer:
```{sql}
SELECT EmployeeId, Id, OrderDate
FROM 'Orders'
WHERE OrderDate = DATE(OrderDate, 'start of month', '+1 month', '-1 day')
ORDER BY EmployeeId, Id
```

### Q36: Orders with many line items  

**Task:** Show the 10 orders with the most line items, in order of total line items.  
  
Show OrderDetails table
```{sql}
SELECT * FROM 'OrderDetail'
```
Show Orders table
```{sql}
SELECT * FROM 'Orders'
```
Final answer:
```{sql}
SELECT Orders.Id, COUNT(*) AS 'TotalOrderDetail'
FROM 'Orders'
LEFT JOIN 'OrderDetail' ON OrderDetail.OrderId = Orders.Id
GROUP BY OrderDetail.OrderId
ORDER BY COUNT(OrderDetail.OrderId) DESC, Orders.Id DESC
LIMIT 10
```

