---
title: 'SQL Practice: Intermediate problems'
output: html_document
---

```{r setup, include=FALSE}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), "db/Northwind_large.sqlite")
knitr::opts_chunk$set(echo = TRUE, connection = "con", max.print = NA)
```

# IMPORTANT!! Rename the 'Order' table into 'Orders'

```{sql}
--ALTER TABLE 'Order' RENAME TO 'Orders'
```

## R Markdown  
Here I'll work with part 1 of "sql practice problems" book


## Connect to DB and overview tables
```{r}
dbListTables(con)
```

### Q20: Categories, and the total products in each category

**Product** table overview:
```{sql}
SELECT * FROM 'Product'
```

For this problem, we’d like to see the total number of products in each category. Sort the results by the total number of products, in descending order
```{sql}
SELECT CategoryId, COUNT(CategoryId) AS 'CategoryNo'
FROM 'Product'
GROUP BY CategoryId
ORDER BY CategoryNo DESC
```
Join **Category** table
```{sql}
SELECT Category.CategoryName, COUNT(Product.CategoryId) AS 'TotalProducts'
FROM 'Product'
JOIN 'Category' ON Product.CategoryId = Category.Id
GROUP BY CategoryName
ORDER BY TotalProducts DESC 
```

### Q21: Total customers per country/city

Overvies **Customer** table
```{sql}
SELECT Id, Country, City FROM 'Customer'
```

In the **Customer** table, show the total number of customers per Country and City.
```{sql}
SELECT Country, City, COUNT(*) AS 'TotalCustomer'
FROM 'Customer' 
GROUP BY Country, City
ORDER BY TotalCustomer DESC
```

### Q22: Products that need reordering  

What products do we have in our inventory that should be reordered?
```{sql}
SELECT * FROM 'Product'
```
For now, just use the fields **UnitsInStock** and **ReorderLevel**, where UnitsInStock is less than the ReorderLevel, ignoring the fields UnitsOnOrder and Discontinued. Order the results by ProductID.
```{sql}
SELECT Id, UnitsInStock, ReorderLevel
FROM 'Product'
WHERE UnitsInStock < ReorderLevel
ORDER BY Id
```

### Q23: Products that need reordering, continued

Now we need to incorporate these fields — UnitsInStock, UnitsOnOrder, ReorderLevel, Discontinued — into our calculation. We’ll define “products that need reordering” with the following:  
- UnitsInStock plus UnitsOnOrder are less than or equal to ReorderLevel
- The Discontinued flag is false (0)

```{sql}
SELECT Id, ProductName, UnitsInStock, UnitsOnOrder, ReorderLevel, Discontinued
FROM 'Product'
WHERE (UnitsInStock + UnitsOnOrder) <= ReorderLevel AND Discontinued = 0
```

### Q24: Customer list by region  

Overview **Customer** table: 
```{sql}
SELECT * FROM 'Customer'
```

A salesperson for Northwind is going on a business trip to visit customers, and would like to see a list of all customers, sorted by region, alphabetically. 
```{sql}
SELECT Id, CompanyName, Region
FROM 'Customer' 
ORDER BY Region ASC
```

However, he wants the customers with no region (null in the Region field) to be at the end, instead of at the top, where you’d normally find the null values.  
Within the same region, companies should be sorted by **CustomerID**.
```{sql}
SELECT Id, CompanyName, Region
FROM 'Customer' 
ORDER BY
  CASE WHEN Region IS NULL THEN 1 ELSE 0 END,
Region, Id ASC
```
### Q25: High freight charges  
Some of the countries we ship to have very high freight charges. We'd like to investigate some more shipping options for our customers, to be
able to offer them lower freight charges.
```{sql}
SELECT Id, OrderDate, Freight, ShipCountry FROM 'Orders'
```

Return the three ship countries with the highest average freight overall, in descending order by average freight.
```{sql}
SELECT ShipCountry, AVG(Freight) AS 'AverageFreight'
FROM 'Orders'
GROUP BY ShipCountry
ORDER BY AverageFreight DESC
LIMIT(3)
```

### Q26: High freight charges - 2015

Let's overview years in **Orders** table
```{sql}
SELECT strftime('%Y', OrderDate) AS 'OrderYear', count(OrderDate)
FROM 'Orders'
GROUP BY OrderYear
```

Now, instead of using all the orders we have, we only want to see orders from the year 2015
```{sql}
SELECT ShipCountry, AVG(Freight) AS 'AverageFreight'
FROM 'Orders'
WHERE strftime('%Y', OrderDate) = '2015'
GROUP BY ShipCountry
ORDER BY AverageFreight DESC
--LIMIT(3)
```

### Q27: High freight charges with between  

```{sql}
SELECT ShipCountry, AVG(Freight) AS 'AverageFreight'
FROM 'Orders'
WHERE OrderDate BETWEEN '2015-01-01' AND '2015-12-31'
GROUP BY ShipCountry
ORDER BY AverageFreight DESC
--LIMIT(3)
```

All orders for 12-31-2015
```{sql}
SELECT Id, ShipCountry, Freight, OrderDate --strftime('%Y-%m-%d-%H-%M-%S', OrderDate)
FROM 'Orders' 
WHERE strftime('%Y-%m-%d', OrderDate) = '2015-12-31'
```

How many orders France has in 2015:
```{sql}
SELECT ShipCountry, Freight, OrderDate--strftime('%Y-%m-%d', OrderDate)
FROM 'Orders' 
WHERE strftime('%Y', OrderDate) = '2015' AND ShipCountry = 'France'
ORDER BY strftime('%Y-%m-%d', OrderDate) DESC
```

### Q28: High freight charges - last year  

We're continuing to work on high freight charges. We now want to get the three ship countries with the highest average freight charges.
```{sql}
SELECT ShipCountry, AVG(Freight) AS 'AverageFreight'
FROM 'Orders'
GROUP BY ShipCountry
ORDER BY AverageFreight DESC
LIMIT(3)
```

But instead of filtering for a particular year, we want to use the last 12 months of order data, using as the end date the last OrderDate in Orders. 
```{sql}

```

### Q29: Inventory list

We're doing inventory, and need to show information like the below, for all orders. Sort by OrderID and Product ID.  
We want to get next columns: EmployeeID, LastName, OrderID, ProductName, Quantity.  
```{sql}
SELECT * FROM 'Orders'
```
  
EmployeeID and LastName we get from **Employee** table:
```{sql}
SELECT * FROM 'Employee'
```
Next, OrderId and Quantity we get from **OrderDetail** table:
```{sql}
SELECT * FROM 'OrderDetail'
```
And finally, ProductName we get from **Product**
```{sql}
SELECT * FROM 'Product'
```

JOIN 'Category' ON Product.CategoryId = Category.Id
Step 1. Join **OrderDetails** table with Quantity to **Orders**
```{sql}
SELECT Orders.Id, Orders.EmployeeId, OrderDetail.Quantity
FROM 'Orders'
JOIN 'OrderDetail' ON Orders.Id = OrderDetail.OrderId
```
Step 2. Join **OrderDetails** table with Quantity to **Orders**
```{sql}
SELECT Orders.Id, Orders.EmployeeId, OrderDetail.Quantity, Employee.LastName, Product.ProductName
FROM 'Orders'
JOIN 'OrderDetail' ON Orders.Id = OrderDetail.OrderId
JOIN 'Employee' ON Orders.EmployeeId = Employee.Id
JOIN 'Product' ON OrderDetail.ProductId = Product.Id
ORDER BY Orders.Id, Product.Id
```

### Q29: Customers with no orders  

There are some customers who have never actually placed an order. Show these customers

First, look into **Customer** table
```{sql}
SELECT * FROM 'Customer'
```

Second, overview **Orders** table:
```{sql}
SELECT * FROM 'Orders'
```
Let's make left join
```{sql}
SELECT Customer.Id AS Customer_CustomersId, Orders.CustomerId AS Orders_CustomerId
FROM 'Customer'
LEFT JOIN 'Orders' ON Orders.CustomerId = Customer.Id
```

This is a good start. It shows all records from the Customers table, and the matching records from the Orders table. However, we only want those records where the CustomerID in Orders is null. You still need a filter.  
Answer from book:
```{sql}
SELECT Customer.Id AS Customer_CustomersId, Orders.CustomerId AS Orders_CustomerId
FROM 'Customer'
LEFT JOIN 'Orders' ON Orders.CustomerId = Customer.Id
WHERE Customer.Id IS NULL
```
Check answer from book:
```{sql}
SELECT CustomerId
FROM 'Orders'
GROUP BY CustomerId
```
So, we have 95 unique rows in **Order** table vs 91 in **Customer**. Let's find these 4 customers  

My answer:
```{sql}
SELECT Customer.Id AS Customer_CustomersId, Orders.CustomerId AS Orders_CustomerId
FROM 'Orders'
LEFT JOIN 'Customer' ON Customer.Id = Orders.CustomerId
WHERE Customer.Id IS NULL
GROUP BY Orders_CustomerId
```

```{sql}
SELECT Customer.Id AS Customer_CustomersId, Orders.CustomerId AS Orders_CustomerId
FROM 'Orders'
LEFT JOIN 'Customer' ON Customer.Id = Orders.CustomerId
WHERE Customer.Id IS NULL
--GROUP BY Orders_CustomerId
```

### Q31: Customers with no orders for EmployeeID 4

One employee (Margaret Peacock, EmployeeID 4) has placed the most orders...
Check it
```{sql}
SELECT EmployeeId, COUNT(EmployeeId) AS 'NumberOfOrdred'
FROM 'Orders'
GROUP BY EmployeeId
ORDER BY NumberOfOrdred DESC
```

... However, there are some customers who've never placed an order with her. Show only those customers who have never placed an order with her.  
Step 1. filter orders with EmployeeId = 4:
```{sql}
SELECT * FROM 'Orders'
WHERE EmployeeId = 4
```

```{sql}
SELECT Customer.Id, Orders.CustomerID
FROM 'Customer'
LEFT JOIN 'Orders' ON Customer.Id = Orders.CustomerId AND Orders.EmployeeId = 4
GROUP BY Customer.Id
--WHERE Orders.CustomerId IS NULL 
```

```{sql}
SELECT Customer.Id, Orders.CustomerID
FROM 'Customer'
LEFT JOIN 'Orders' ON Customer.Id = Orders.CustomerId AND Orders.EmployeeId = 4
WHERE Orders.CustomerId IS NULL 
```

