---
title: 'SQL Practice: Intermediate problems'
output: html_document
---

```{r setup, include=FALSE}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), "db/Northwind.sqlite")
knitr::opts_chunk$set(echo = TRUE, connection = "con")
```

## R Markdown  
Here I'll work with part 1 of "sql practice problems" book

## Connect to DB and overview tables
```{r }
dbListTables(con)
```

### Q20: Categories, and the total products in each category

**Product** table overview:
```{sql}
SELECT * FROM 'Product'
```

For this problem, we’d like to see the total number of products in each category. Sort the results by the total number of products, in descending order
```{sql}
SELECT CategoryId, COUNT(CategoryId) AS 'CategoryNo'
FROM 'Product'
GROUP BY CategoryId
ORDER BY CategoryNo DESC
```
Join **Category** table
```{sql}
SELECT Category.CategoryName, COUNT(Product.CategoryId) AS 'TotalProducts'
FROM 'Product'
JOIN 'Category' ON Product.CategoryId = Category.Id
GROUP BY CategoryName
ORDER BY TotalProducts DESC 
```

### Q21: Total customers per country/city

Overvies **Customer** table
```{sql}
SELECT Id, Country, City FROM 'Customer'
```

In the **Customer** table, show the total number of customers per Country and City.
```{sql}
SELECT Country, City, COUNT(*) AS 'TotalCustomer'
FROM 'Customer' 
GROUP BY Country, City
ORDER BY TotalCustomer DESC
```

### Q22: Products that need reordering  

What products do we have in our inventory that should be reordered?
```{sql}
SELECT * FROM 'Product'
```
For now, just use the fields **UnitsInStock** and **ReorderLevel**, where UnitsInStock is less than the ReorderLevel, ignoring the fields UnitsOnOrder and Discontinued. Order the results by ProductID.
```{sql}
SELECT Id, UnitsInStock, ReorderLevel
FROM 'Product'
WHERE UnitsInStock < ReorderLevel
ORDER BY Id
```

### Q23: Products that need reordering, continued

Now we need to incorporate these fields — UnitsInStock, UnitsOnOrder, ReorderLevel, Discontinued — into our calculation. We’ll define “products that need reordering” with the following:  
- UnitsInStock plus UnitsOnOrder are less than or equal to ReorderLevel
- The Discontinued flag is false (0)
```{sql}
SELECT Id, ProductName, UnitsInStock, UnitsOnOrder, ReorderLevel, Discontinued
FROM 'Product'
WHERE (UnitsInStock + UnitsOnOrder) <= ReorderLevel AND Discontinued = 0
```

### Q24: Customer list by region  

Overview **Customer** table: 
```{sql}
SELECT * FROM 'Customer'
```

A salesperson for Northwind is going on a business trip to visit customers, and would like to see a list of all customers, sorted by region, alphabetically. 
```{sql}
SELECT Id, CompanyName, Region
FROM 'Customer' 
ORDER BY Region ASC
```

However, he wants the customers with no region (null in the Region field) to be at the end, instead of at the top, where you’d normally find the null values.  
Within the same region, companies should be sorted by **CustomerID**.
```{sql}
SELECT Id, CompanyName, Region
FROM 'Customer' 
ORDER BY
  CASE WHEN Region IS NULL THEN 1 ELSE 0 END,
Region, Id ASC
```
### Q25: High freight charges  
Some of the countries we ship to have very high freight charges. We'd like to investigate some more shipping options for our customers, to be
able to offer them lower freight charges.
```{sql}
SELECT Id, OrderDate, Freight, ShipCountry FROM 'Orders'
```

Return the three ship countries with the highest average freight overall, in descending order by average freight.
```{sql}
SELECT ShipCountry, AVG(Freight) AS 'AverageFreight'
FROM 'Orders'
GROUP BY ShipCountry
ORDER BY AverageFreight DESC
LIMIT(3)
```








