---
title: 'SQL: Introductary Problems'
output: html_document
---

```{r setup, include=FALSE}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), "db/Northwind.sqlite")
knitr::opts_chunk$set(echo = TRUE, connection = "con")
```

## R Markdown  
Here I'll work with part 1 of "sql practice problems" book

## Connect to DB and overview tables
```{r }
dbListTables(con)
```

### Q1: Which shippers do we have?  

We have a table called Shippers. Return all the fields from all the shippers
```{sql, connection=con}
SELECT * FROM 'Shipper'
```

### Q:How to count number of rows
```{sql, connection=con}
SELECT COUNT(*) FROM 'Customer'
```

### Q2: Select certain fields from table
In the Categories table, selecting all the fields using this SQL
```{sql, connection=con}
SELECT * FROM 'Category'
```

... we only want 2 columns CategoryName and Description
```{sql, connection=con}
SELECT CategoryName, Description FROM 'Category'
```

### Q3: We’d like to see just the FirstName, LastName, and HireDate of all the employees with the Title of Sales Representative  

Overvies data
```{sql, connection=con}
SELECT * FROM 'Employee'
```

Answer the question
```{sql, connection=con}
SELECT FirstName, HireDate
FROM 'Employee'
WHERE Title = 'Sales Representative'
```
### Q4: Sales Representatives in the United States  

Now we’d like to see the same columns as above, but only for those employees that both have the title of Sales Representative, and also are in the United States.  
```{sql, connection=con}
SELECT FirstName, HireDate
FROM 'Employee'
WHERE Title = 'Sales Representative' AND Country = 'USA'
```

### Q5: Orders placed by specific EmployeeID  

Order table overview
```{sql}
SELECT * FROM 'Order'
```

Show all the orders placed by a specific employee The EmployeeID for this Employee (Steven Buchanan) is 5.
```{sql}
SELECT Id AS 'OrderID', OrderDate FROM 'Order' WHERE EmployeeId = 5
```

### Q6: Suppliers and ContactTitles

```{sql}
SELECT * FROM 'Supplier'
```

In the Suppliers table, show the SupplierID, ContactName, and ContactTitle for those Suppliers whose ContactTitle is not Marketing Manager
```{sql}
SELECT ID AS SupplierID, ContactName, ContactTitle
FROM 'Supplier' 
WHERE ContactTitle != 'Marketing Manager'
```

### Q7: Products with “queso” in ProductName
Overview Product table:
```{sql}
SELECT * FROM 'Product'
```

In the **Product** table, we’d like to see the **ProductID** and **ProductName** for those products where the **ProductName** includes the string “queso”
```{sql}
SELECT Id AS ProductID, ProductName
FROM Product
WHERE ProductName LIKE "%queso%"
```
### Q8: Orders shipping to France or Belgium

Looking at the **Order** table
```{sql}
SELECT * FROM 'Order'
```

Looking at the **Order** table, there’s a field called **ShipCountry**. Write a query that shows the **OrderID**, **CustomerID**, and **ShipCountry** for the orders where the **ShipCountry** is either France or Belgium.
```{sql}
SELECT Id AS OrderID, CustomerId, ShipCountry
FROM 'Order'
WHERE ShipCountry IN ('France', 'Belgium')
```

### Q9: Orders shipping to any country in Latin America  

Now, instead of just wanting to return all the orders from France of Belgium, we want to show all the orders from any Latin American country
```{sql}
SELECT Id AS OrderID, CustomerId, ShipCountry
FROM 'Order'
WHERE ShipCountry IN ('Brazil', 'Mexico', 'Argentina', 'Venezuela')
```

### Q10: Employees, in order of age  

Look into **Employee** table
```{sql}
SELECT * FROM 'Employee'
```

For all the employees in the **Employee** table, show the **FirstName**, **LastName**, **Title**, and **BirthDate**. Order the results by **BirthDate**, so we have the oldest employees first.
```{sql}
SELECT FirstName, LastName, Title, BirthDate
FROM 'Employee' 
ORDER BY BirthDate
```

### Q12: Employees full name

Show the **FirstName** and **LastName** columns from the **Employees** table, and then create a new column called FullName, showing FirstName and LastName joined together in one column, with a space in-between
```{sql}
SELECT FirstName, LastName, FirstName || ' ' || LastName AS FullName
FROM 'Employee'
```

### Q13: OrderDetails amount per line item  

In the **OrderDetails** table, we have the fields **UnitPrice** and **Quantity**
```{sql}
SELECT * FROM 'OrderDetail'
```

Create a new field, **TotalPrice**, that multiplies these two together. We’ll ignore the Discount field for now.  
In addition, show the **OrderID**, **ProductID**, **UnitPrice**, and **Quantity**. Order by **OrderID** and **ProductID**.
```{sql}
SELECT OrderID, ProductID, UnitPrice, Quantity, UnitPrice*Quantity AS TotalPrice
FROM 'OrderDetail'
ORDER BY OrderId, ProductID
```

### Q14: How many customers?

How many customers do we have in the Customers table?
```{sql}
SELECT * FROM 'Customer'
```

Show one value only, and don’t rely on getting the recordcount at the end of a resultset.
```{sql}
SELECT COUNT(*) AS 'Number of customers' FROM 'Customer'
```

### Q15: When was the first order?

Overview **Order** table
```{sql}
SELECT * FROM 'Order'
```

Show the date of the first order ever made in the **Order** table
```{sql}
SELECT OrderDate AS 'First order date'
FROM 'Order'
ORDER BY OrderDate ASC
LIMIT 1
```

### Q16: Countries where there are customers  

Show a list of countries ...
```{sql}
SELECT * FROM 'Customer'
```

... where the Northwind company has customers
```{sql}
SELECT Country FROM 'Customer' GROUP BY Country
```

### Q17: Contact titles for customers

Show a list of all the different values in the **Customer** table for **ContactTitle**. Also include a count for each **ContactTitle**
```{sql}
SELECT ContactTitle, COUNT(ContactTitle) AS 'CountTitle'
FROM 'Customer'
GROUP BY ContactTitle
ORDER BY CountTitle DESC
```

### Q18: Products with associated supplier names

We’d like to show, for each product, the associated Supplier. 
```{sql}
SELECT * FROM 'Product'
```
**Supplier** table overview
```{sql}
SELECT * FROM 'Supplier'
```

Show the **ProductID**, **ProductName**, and the **CompanyName** of the Supplier. Sort by **ProductID**.  
```{sql}
SELECT Product.Id, ProductName, CompanyName
FROM 'Product'
JOIN 'Supplier' ON Product.SupplierId = Supplier.Id 
```

### Q19: Orders and the Shipper that was used

We’d like to show a list of the Orders that were made, including the Shipper that was used.
```{sql}
SELECT * FROM 'Orders'
```
Overview Shipper table:
```{sql}
SELECT * FROM 'Shipper'
```

Show the OrderID, OrderDate (date only), and CompanyName of the Shipper, and sort by OrderID.
```{sql}
SELECT `Order.Id`, OrderDate, CompanyName
FROM 'Order'
JOIN 'Shipper' ON `Order.ShipVia` = Shipper.Id
```

Have some problem with table Order's name :( Let's temporary rename **Order** table into **Orders** table.  
The command is: ALTER TABLE 'Order' RENAME TO 'Orders'  
I've used for this [sqllite simulator](https://sqliteonline.com/)  
  
In order to not show all the orders (there’s more than 800), show only those rows with an OrderID of less than 10300.
```{sql}
SELECT Orders.Id, OrderDate, CompanyName
FROM 'Orders'
JOIN 'Shipper' ON Orders.ShipVia = Shipper.Id
WHERE Orders.ID < 10300
```

