---
title: "Untitled"
output: html_document
---

# Here is we convert sql tables to r-objects

```{r setup, include=FALSE}
library(RSQLite)
library(tidyverse)
library(lubridate)

sqllarge <- dbConnect(RSQLite::SQLite(), "db/Northwind_large.sqlite")
knitr::opts_chunk$set(echo = TRUE, connection = "sqllarge")
```

## Connect to DB and overview tables
```{r}
dbListTables(sqllarge)
```

## Convert sql-tables to R-objects

R-object customers:
```{sql, output.var = 'customers'}
SELECT * FROM 'Customer'  
```
View and save r-object to file
```{r}
customers
#save(customers, file = 'db/customers_table.Rdata')
```
R-object order_details:
```{sql, output.var = 'order_details'}
SELECT * FROM 'OrderDetail'
```
View and save r-object to file:
```{r}
order_details
#save(order_details, file = 'db/order_details_table.Rdata')
```
R-object orders:
```{sql, output.var = 'orders'}
SELECT * FROM 'Orders'
```
View and save r-object to file:
```{r}
orders
```
First, transform OrderDate columb
```{r}
orders$OrderDate[c(1, 7500)]
```
So, we have problem: different format to dates. Let's correct it:
```{r}
orders$OrderDate[c(1, 7500)] %>% 
  ymd_hms(truncated = 3)
```
Now let's try to convert dates from orders dataframe:
```{r}
orders$OrderDate <- ymd_hms(orders$OrderDate, truncated = 3)
class(orders$OrderDate)
```
Save the dataframe with corrected date's type
```{r, eval=FALSE}
save(orders, file = 'db/orders_table.Rdata')
```

