---
title: "Untitled"
output: html_document
---

# Here is we convert sql tables to r-objects

```{r setup, include=FALSE}
library(RSQLite)
library(tidyverse)
library(lubridate)

sqllarge <- dbConnect(RSQLite::SQLite(), "db/Northwind_large.sqlite")
sqlsmall <- dbConnect(RSQLite::SQLite(), "db/edit_Northwind.sqlite")

knitr::opts_chunk$set(echo = TRUE)
```

## Connect to DB and overview tables
```{r}
dbListTables(sqllarge)
```

## Convert sql-tables to R-objects

### Customers table - Large db  
```{sql, connection = 'sqllarge', output.var = 'customers'}
SELECT * FROM 'Customer'  
```
View and save r-object to file
```{r}
customers
#save(customers, file = 'db/customers_table.Rdata')
```

### Customers table - Small db  

```{sql, connection = 'sqlsmall', output.var = 'small_customer'}
SELECT * FROM 'Customer'  
```
View and save r-object to file
```{r}
small_customer
#save(small_customer, file = 'db/small_customer_table.Rdata')
```
### OrderDetail table - Large db  
  
R-object order_details:
```{sql, connection = 'sqllarge', output.var = 'order_details'}
SELECT * FROM 'OrderDetail'
```
View and save r-object to file:
```{r}
order_details
#save(order_details, file = 'db/order_details_table.Rdata')
```

### OrderDetail table - Small db  
  
R-object order_details:
```{sql, connection = 'sqlsmall', output.var = 'small_order_details'}
SELECT * FROM 'OrderDetail'
```
View and save r-object to file:
```{r}
small_order_details
#save(small_order_details, file = 'db/small_order_details_table.Rdata')
```
### Orders table - Large db  

R-object orders:
```{sql, connection = 'sqllarge', output.var = 'orders'}
SELECT * FROM 'Orders'
```
View and save r-object to file:
```{r}
orders
```
First, transform OrderDate columb
```{r}
orders$OrderDate[c(1, 7500)]
```
So, we have problem: different format to dates. Let's correct it:
```{r}
orders$OrderDate[c(1, 7500)] %>% 
  ymd_hms(truncated = 3)
```
Now let's try to convert dates from orders dataframe:
```{r}
orders$OrderDate <- ymd_hms(orders$OrderDate, truncated = 3)
class(orders$OrderDate)
```
Save the dataframe with corrected date's type
```{r, eval=FALSE}
#save(orders, file = 'db/orders_table.Rdata')
```

### Orders table - Small db  

R-object orders:
```{sql, connection = 'sqlsmall', output.var = 'small_orders'}
SELECT * FROM 'Orders'
--strftime('%Y', OrderDate) = '2015'
```
View and save r-object to file:
```{r}
small_orders
#save(small_orders, file = 'db/small_orders_table.Rdata')
```

## Difference questions  

### Compare OrderDate column in two Orders tables from large and small dbs:  
```{r}
left_join(small_orders %>% 
            select('sId' = Id, 'sOrderDate' = OrderDate),
          orders %>% 
            select(Id, OrderDate),
          by = c('sId' = 'Id')) %>% 
  arrange(OrderDate)
```
What about years:
```{r}
# Run previous chunk!!!
count(.Last.value, 'Year' = year(OrderDate))
```
Edit OrderDate table in small db with next replaces:
1996 -> 2014
1997 -> 2015
1998 -> 2016

DONE in edit_Northwind.sql table!!! 

